The Rockerverse: Packages and Applications for Containerisation with R

### 2020-03-02 manuscript.Rmd
one sentence per line, slight shortening

First, different packages use containers for the **deployment of R and \CRANpkg{Shiny} apps**.
\CRANpkg{Shiny} is a popular package for creating interactive online dashboards with R, and it enables users with very diverse backgrounds to create stable and user-friendly web applications.
_ShinyProxy_ ([https://www.shinyproxy.io/](https://www.shinyproxy.io/)) is an open source tool to deploy Shiny apps in an enterprise context, where it features single sign-on, but it can also be used in scientific use cases \citep[e.g., ][]{savini_epiexplorer_2019,glouzon_structurexplor_2017}.
ShinyProxy uses Docker containers to isolate user sessions and to achieve scalability for multi-user scenarious with multiple apps.
ShinyProxy itself is written in Java to accomodate corporate requirements and may itself run in a container for stability and availability.
The tool is built on ContainerProxy ([https://www.containerproxy.io/](https://www.containerproxy.io/)), which provides similar features for executing long-running R jobs or interactive R sessions.
The started containers can run on a regular Docker host but also in clusters.
Continuous integration and deployment (CI/CD) for Shiny applications using Shinyproxy on virtual machines
can be achieved e.g. via [GitLab pipelines](https://docs.gitlab.com/ee/ci/pipelines.html). 
A pipeline can include building and checking R packages and Shiny apps. After the code has passed
checks, Docker images are built and pushed to GitLab's container registry
(public or private, depending on project settings). The pipeline is finished off by sending an HTTP
request to the virtual machine where and endpoint (webhook) is waiting to execute the
deployment script (updating configurations, pulling new Docker images, restarting services).

@@

Parallel efforts to enable R for different cloud providers leads to some level of code fragmentation. This fact and to avoid lock-in motivated the [OpenFaaS project](https://www.openfaas.com/), that facilitates the deployment of functions and microservices to Kubernetes or Docker Swarm. OpenFaaS is language agnostic, it provides auto-scaling, metrics, and an API gateway.
Reduced boilerplate code is achieved via templates. R templates\footnote{OpenFaaS R templates: \href{https://github.com/analythium/openfaas-rstats-templates}{https://github.com/analythium/openfaas-rstats-templates}}
are provided based on Rocker's Debian and R-hub's [r-minimal](https://github.com/r-hub/r-minimal) Alpine images.
The templates use [multi-stage Docker builds](https://docs.docker.com/develop/develop-images/multistage-build/) to combine R base images with the OpenFaaS 'watchdog', a tiny Golang webserver that marshals an HTTP request and invokes the application. 
The R session uses \CRANpkg{plumber} or similar packages for the API endpoint with packages and data preloaded, thus minimizing response times.

>>>>

First, different packages use containers for the **deployment of R and \CRANpkg{Shiny} apps**.
\CRANpkg{Shiny} is a popular package for creating interactive online dashboards with R, and it enables users with very diverse backgrounds to create stable and user-friendly web applications.
_ShinyProxy_ ([https://www.shinyproxy.io/](https://www.shinyproxy.io/)) is an open source tool to deploy Shiny apps in an enterprise context, where it features single sign-on, but it can also be used in scientific use cases \citep[e.g., ][]{savini_epiexplorer_2019,glouzon_structurexplor_2017}.
ShinyProxy uses Docker containers to isolate user sessions and to achieve scalability for multi-user scenarious with multiple apps.
ShinyProxy itself is written in Java to accomodate corporate requirements and may itself run in a container for stability and availability.
The tool is built on ContainerProxy ([https://www.containerproxy.io/](https://www.containerproxy.io/)), which provides similar features for executing long-running R jobs or interactive R sessions.
The started containers can run on a regular Docker host but also in clusters.
Continuous integration and deployment (CI/CD) for Shiny applications using Shinyproxy can be achieved, e.g., via [GitLab pipelines](https://docs.gitlab.com/ee/ci/pipelines.html).
A pipeline can include building and checking R packages and Shiny apps.
After the code has passed the checks, Docker images are built and pushed to GitLab's container registry.
The pipeline finishes with triggering a webhook on the server, where the
deployment script is executed.
The script updates configurations, pulls the new Docker images, and restarts the services.

@@

Independent integrations of R for different cloud providers lead to repeated efforts and code fragmentation.
To mitigate these problems and to avoid vendor lock-in motivated the [OpenFaaS project](https://www.openfaas.com/).
OpenFaas facilitates the deployment of functions and microservices to Kubernetes or Docker Swarm.
It is language agnostic and provides auto-scaling, metrics, and an API gateway.
Reduced boilerplate code is achieved via templates.
Template for R\footnote{See OpenFaaS R templates at \href{https://github.com/analythium/openfaas-rstats-templates}{https://github.com/analythium/openfaas-rstats-templates}.}
are provided based on Rocker's Debian and R-hub's [r-minimal](https://github.com/r-hub/r-minimal) Alpine images.
The templates use [multi-stage Docker builds](https://docs.docker.com/develop/develop-images/multistage-build/) to combine R base images with the OpenFaaS 'watchdog', a tiny Golang webserver.
The watchdog marshals an HTTP request and invokes the actual application.
The R session uses \CRANpkg{plumber} or similar packages for the API endpoint with packages and data preloaded, thus minimizing response times.
